---
title: "Geffo"
author: "Olushina A. Junior Ale"
date: "`r Sys.Date()`"

#output: html_document

output:
  rmdformats::html_clean:
    self_contained: yes
    thumbnails: no
    lightbox: yes
    gallery: no
    code_folding: show
    download_code: yes
    number_sections: true
    css: style.css
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  out.width = "100%",
  out.height = "50%",
  fig.align = "center"
  )

```
 
```{r loading libraries, include=FALSE}

pacman::p_load(
    dplyr,
    readxl,
    ggplot2,
    janitor,
    skimr,
    plotly,
    tidyr,
    ggpubr,
    ggrepel,
    flextable,
    formattable,
    gtsummary,
    knitr,
    kableExtra,
    tibble,
    gt,
    stringr
    )

```

# **1.	Part 1: Data Cleaning**
## **Load the dataset and check the structure:**

```{r 0, echo = FALSE}

data <- read_excel("malnutrition_estimates.xlsx")

# Check structure of the dataset
str(data)

```
## **b.	Identify the columns with missing values. How many missing values are in the severe wasting column?**

```{r 1, echo=FALSE}

# Check for missing values in all columns
colSums(is.na(data))

# Count missing values in the 'Severe Wasting' column
sum(is.na(data$`Severe Wasting`))

```
## **Impute missing values in the selected columns with their respective means:**

```{r 2, echo=FALSE}

# Impute missing values with column means
cols_to_impute <- c('Severe Wasting', 'Wasting', 'Overweight', 'Stunting', 'Underweight')

for (col in cols_to_impute) {
  data[[col]][is.na(data[[col]])] <- mean(data[[col]], na.rm = TRUE)
}

```

## **Convert variables into factors:**

```{r 3, echo=FALSE}

# Convert necessary columns into factors

data$`Income Classification` <- as.factor(data$`Income Classification`)
data$LDC <- as.factor(data$LDC)
data$LIFD <- as.factor(data$LIFD)
data$`LLDC or SIDS` <- as.factor(data$`LLDC or SIDS`)

```

## **Remove non-essential variables:**

```{r 4, echo=FALSE}
# Remove unnecessary columns
data <- data[, !names(data) %in% c('ISO code', 'Notes', 'Report Author', 'Source', 'Short Source')]

```

## **Save the cleaned dataset:**

```{r 5, echo=FALSE}
# Save cleaned dataset
write.csv(data, "mal_clean.csv", row.names = FALSE)

```

# **2.	Part 2: Data Analysis**

## **Create a correlation matrix and interpret relationships:**

```{r 6, echo=FALSE}
# Compute the correlation matrix for numeric variables
num_data <- data[, sapply(data, is.numeric)]
corr_matrix <- cor(num_data, use = "complete.obs")
print(corr_matrix)
```

**Interpretation:**

## **Linear regression analysis:**

```{r 7, echo=FALSE}

# Linear regression: Underweight as a function of other variables
lm_model <- lm(Underweight ~ `Severe Wasting` + Wasting + Overweight + Stunting + Year + `U5 Population ('000s)`, data = data)
summary(lm_model)

```
**Is the model a good fit for predicting underweight prevalence? Why or why not?**

**What is the estimated effect of “overweight” on “underweight”, holding other variables constant?**

## **Identify top 10 countries with the highest "overweight" values:**

```{r 8, echo=FALSE}
# Sort data by Overweight and get top 10
top_10_overweight <- data[order(-data$Overweight), c("Country", "Year", "Overweight")]
top_10_overweight <- head(top_10_overweight, 10)
print(top_10_overweight)

```

# **Part 3: Data Visualization in R**

## **Scatter plot of "Wasting" vs "Stunting" with color by "LLDC or SIDS":**

```{r 9, echo=FALSE}

# Scatter plot
ggplot(data, aes(x = Stunting, y = Wasting, color = `LLDC or SIDS`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship between Wasting and Stunting",
       x = "Stunting (%)",
       y = "Wasting (%)") +
  theme_minimal() +
    theme(panel.background = element_rect(fill = 'white'),
          panel.grid = element_blank(),
          axis.line = element_line(color = 'black'),
          text = element_text(size = 12))

ggsave("Scatter plot.png", width = 8, height = 6)

```

## **Box plot of "Overweight" by "Income Classification":**

```{r 10, echo=FALSE}

ggplot(data, aes(x = `Income Classification`, y = Overweight)) +
  geom_boxplot() +
  labs(title = "Distribution of Overweight by Income Classification",
       x = "Income Classification",
       y = "Overweight (%)") +
  theme_minimal() +
    theme(panel.background = element_rect(fill = 'white'),
          panel.grid = element_blank(),
          axis.line = element_line(color = 'black'),
          text = element_text(size = 12))

ggsave("Box plot.png", width = 8, height = 6)
```

## **Histogram of "Underweight" and interpret skewness:**

```{r 11, echo=FALSE}

# Histogram
ggplot(data, aes(x = Underweight)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  labs(title = "Distribution of Underweight",
       x = "Underweight (%)",
       y = "Frequency") +
  theme_minimal() +
    theme(panel.background = element_rect(fill = 'white'),
          panel.grid = element_blank(),
          axis.line = element_line(color = 'black'),
          text = element_text(size = 12))

ggsave("Histogram.png", width = 8, height = 6)
```


```{r 12, echo=FALSE}
```


```{r 13, echo=FALSE}
```


```{r 14, echo=FALSE}
```


```{r 15, echo=FALSE}
```


```{r 16, echo=FALSE}
```


```{r 17, echo=FALSE}
```


```{r 18, echo=FALSE}
```


```{r 19, echo=FALSE}
```
'
```{r 20, echo=FALSE}
```

